Alignment of Light and Mass in Lensing Galaxies

### 2013-05-07 shapes/paper.tex
update on paper

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{myaasmacros}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amsmath}

% Some definitions of things I always use here:
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\simlt{\lower.5ex\hbox{\ltsima}}   
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\simgt{\lower.5ex\hbox{\gtsima}}
\newcommand\bcite[1]{\citeauthor{#1} \citeyear{#1}}

% Some definitions for the priors:
\def\gprior{{\tt gprior}}
\def\cprior{{\tt cprior}}
\def\bprior{{\tt bprior}}
\def\lbprior{{\tt lbprior}}

%Some definitions for rhodm etc:
\def\vztwo{\overline{v_z^2}}
\def\vztwoi{\overline{v_{z,i}^2}}
\def\rhodisc{\rho_\mathrm{disc}(z)}
\def\rhodmext{\rho_\mathrm{dm,ext}}
\def\rhodm{\rho_\mathrm{dm}}
\def\rhoeff{\rho_\mathrm{dm}^\mathrm{eff}}
\def\nuobs{\nu_\mathrm{obs}(z)}

@@

\author[Bruderer]{Claudio Bruderer$^1$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{1,2}$, P. Saha$^3$, J. Coles$^3$\\
$^1$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^2$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^3$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland
}

@@

\begin{abstract}
Abstract here. 
\end{abstract}

\begin{keywords}
\end{keywords}

\section{Introduction}

@@

\section{Data}
\begin{itemize}
\item SPS (Ignacio Ferreras, Dominik, etc. \citep{2011ApJ...740...97L})
\item Write why specifically these lenses were chosen, where they're from
\end{itemize}

\begin{table}
 \begin{center}
  \begin{tabular}{l c c l | l c c l}
   Lens & $z_{L}$ & $z_{S}$ & Type & Lens & $z_{L}$ & $z_{S}$ & Type\\ \hline \hline
   Q0047 & 0.48 & 3.60 & Quad & HE1104 & 0.73 & 2.32 & Double\\
   Q0142 & 0.49 & 2.72 & Double & PG1115 & 0.31 & 1.72 & Quad\\
   MG0414 & 0.96 & 2.64 & Quad & B1152 & 0.44 & 1.02 & Double\\
   B0712 & 0.41 & 1.34 & Quad & B1422 & 0.34 & 3.62 & Quad\\
   RXJ0911 & 0.77 & 2.80 & Quad & B1608 & 0.63 & 1.39 & Quad\\
   BRI0952 & 0.63 & 4.50 & Double & MG2016 & 1.01 & 3.27 & Quad$^{1}$\\
   Q0957 & 0.36 & 1.41 & Double & B2045 & 0.87 & 1.28 & Quad\\
   LBQS1009 & 0.87 & 2.74 & Double & HE2149 & 0.60 & 2.03 & Double\\
   B1030 & 0.60 & 1.54 & Double & Q2237 & 0.04 & 1.69 & Quad\\
%   SBS1520 & 0.72 & 1.86 & Double
%	B1600 & 0.41 & 1.59 & Double\\
  \end{tabular}
  \caption{List of all the lenses in the sample and their properties. Data is taken from \textit{CASTLES} and \cite{leier11phd}. The angles are measured north-through-east. \newline $^{1}$ Two images merge, thus only three images can be resolved.}
  \label{tab:overview}
 \end{center}
\end{table}

\begin{table}
 \begin{center}
  \begin{tabular}{l c c c c c c c}
   \multirow{2}{*}{Lens} & \multirow{2}{*}{Environment$^{a}$} & \multicolumn{2}{c}{Group/Cluster centroid$^{b}$} & & \multicolumn{2}{c}{Other galaxy} & \multirow{2}{*}{References$^{c}$}\\ \cline{3-4} \cline{6-7}
    & & $\Delta\alpha$ [''] & $\Delta\delta$ [''] & & $\Delta\alpha$ [''] & $\Delta\delta$ ['']\\ \hline \hline
   Q0047 & G(9) & $3.6\pm15.6$ & $43.2\pm28.8$ &   & ... & ... & 1\\
   MG0414 & 2 & ... & ... &   & $-0.4$ & $1.5$ & 2\\
   B0712 & 2 & ... & ... &   & $47$ & $-91$ & 3\\
   RXJ0911 & C & $-12.0$ & $-39.8$ &    & ... & ... & 4\\
   Q0957 & C & ... & ... &   & ... & ... & 5\\
   B1030 & 2 & ... & ... &   & $-9.8$ & $8.5$ & 6\\
   PG1115 & G(13) & $-10.8\pm21.0$ & $-3.6\pm15.6$ &   & ... & ... & 1\\
   B1422 & G(17) & $61.2\pm25.2$ & $-10.8\pm23.4$ &   & ... & ... & 1\\
%   SBS1520 & G(5) & ... & ... &   & ... & ... & 7\\
%   B1600 & G(17) & $-8.4^{d}$ & $-1.5^{d}$ &   & ... & ... & 8\\
   B1608 & G(8) & $-0.7$ & $8.0$ &   & ... & ... & 9\\
   MG2016 & C(69) & ... & ... &   & ... & ... & 10\\
   HE2149 & G & $-36.6$ & $14.2$ &   & ... & ... & 11\\
  \end{tabular}
  \caption{List of lenses lying in a group or cluster environment, or that have a companion. \newline $^{a}$ G stands for group, C for cluster, and 2 means that the lens has a companion galaxy. If known, the number of member galaxies including the lens is given. \newline $^{b}$ If not stated differently, the centroid positions are luminosity-weighted estimates. \newline $^{c}$ References for the centroid estimates we used. \newline $^{d}$ Only number-weighted estimates are available. \newline 1: \cite{2011ApJ...726...84W}; 2: \cite{1993AJ....105....1S}; 3: \cite{2002AJ....123..627F}; 4: \cite{2001ApJ...555....1M}; 5: \cite{1998ApJ...504..661C}; 6: \cite{2000ApJ...536..584L}; 7: \cite{2008ApJ...673..778A}; 8: \cite{2007AJ....134..668A}; 9: \cite{2003MNRAS.344..337T}; 10: \cite{2006ApJ...646...85W}; 11: \cite{2006ApJ...646...85W}}
  \label{tab:environment}
 \end{center}
\end{table}

\begin{table}
 \begin{center}
  \begin{tabular}{l c c c c}
   Lens & \multicolumn{3}{c}{Time delays [d]} & References\\ \hline \hline
   RXJ0911 & $\Delta t_{BA}=146\pm4$ & ... & ... & 1\\
   Q0957 & $\Delta t_{BA}=416.5\pm1.0^{a}$ & $\Delta t_{BA}=420.6\pm1.9^{b}$ & ... & 2\\
   HE1104 & $\Delta t_{BA}=162.2^{+6.3}_{-5.9}$ & ...  & ...  & 3\\
   PG1115 & $\Delta t_{BA}=12.0^{+2.4}_{-2.0}$ & $\Delta t_{DC}=4.4^{+3.2}_{-2.4}$ & ... & 4\\
%   SBS1520 & $\Delta t_{BA}=125.8\pm2.1$ & ... & ... & 5\\
%   B1600 & $\Delta t_{BA}=51\pm2$ & ... & ... & 6\\
   B1608 & $\Delta t_{BA}=31.5^{+2}_{-1}$ & $\Delta t_{CA}=36.0\pm1.5$ & $\Delta t_{DA}=77.0^{+2}_{-1}$ & 7\\
   HE2149 & $\Delta t_{BA}=103\pm12$ & ... & ... & 8\\
  \end{tabular}
  \caption{Lenses with measured time delays. $\Delta t_{BA}$ corresponds to the time delay between the images A and B with A leading. \newline $^{a}$ measured in the g-band \newline $^{b}$ measured in the r-band \newline 1: \cite{2002ApJ...572L..11H}; 2: \cite{2012A&A...540A.132S}; 3: \cite{2008ApJ...676...80M}; 4: \cite{2010MNRAS.406.2764T}; 5: \cite{2011A&A...536A..44E}; 6: \cite{2000ApJ...544..117B}; 7: \cite{2002ApJ...581..823F}; 8: \cite{2002A&A...383...71B}}
  \label{tab:timedelays}
 \end{center}
\end{table}

@@

Our sample consists of 20 galaxies. The sample is identical to the one as in \cite{leier11phd} and \cite{2011ApJ...740...97L} except that \textit{HS0818+1227} proved too difficult to reconstruct and therefore was left out. Most of the lens galaxies are early-type galaxies, except \textit{B1030+074}, \textit{B1152+200}, \textit{B1600+434}, \textit{B1608+656}, and \textit{Q2237+030}. All lenses can also be found in the gravitational lens database \textit{CASTLES}. Characteristics and references of each lens in the sample are listed below. Table \ref{tab:overview} contains of all the redshifts and the type of the lenses, Table \ref{tab:environment} lenses with environment information, and Table \ref{tab:timedelays} lenses with measured time delays.

@@

\textit{B1608+656} (hereafter \textit{B1608}) consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment and the mass distribution along the line of sight have been analyzed by \cite{2006ApJ...642...30F}. They find a group with 8 resp. 9 members if the merging galaxies are counted as 2. No significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. Along the line of sight seem to be four other groups.

@@

NOTE: Cannot use B1608, but I still left it in the sample as Fei probably will use the lens.

\section{Methods}
\subsection{GLASS}
\begin{itemize}
\item Explain GLASS briefly and link to fake data tests
\end{itemize}

@@

\section{Results}
\begin{itemize}
\item Show some examples (stellar and dm contour plots, image of lens, arrival time surface, included mass)
\item Show money plots
\item Comment on a few lenses special lenses
\item Environmental effect?
\end{itemize}

\begin{figure}
 \begin{center}
  \hspace{-7mm}\includegraphics[height=0.1\textwidth]{Figures/1115_a.eps}
  \hspace{-7mm}\includegraphics[height=0.1\textwidth]{Figures/1115_b.eps}
  \hspace{-7mm}\includegraphics[height=0.1\textwidth]{Figures/1115_c.eps}
  \hspace{-7mm}\includegraphics[height=0.1\textwidth]{Figures/1115_d.eps}
  \hspace{-7mm}\includegraphics[height=0.1\textwidth]{Figures/1115_h.eps}\\
  \caption{Reconstruction of lens \textit{PG1115}}
  \label{fig:1115rec}
 \end{center}
\end{figure}

\begin{figure}
 \begin{center}
 \hspace{-7mm}\includegraphics[height=0.29\textwidth]{Figures/b.eps}
 \hspace{-7mm}\includegraphics[height=0.29\textwidth]{Figures/d.eps}
 \hspace{-7mm}\includegraphics[height=0.29\textwidth]{Figures/e.eps}
 \caption{Plots of stellar and dark matter misalignments, ellipticities and composed plot.}
 \label{fig:moneyplots}
 \end{center}
\end{figure}

\begin{figure}
 \begin{center}
 \hspace{-7mm}\includegraphics[height=0.29\textwidth]{Figures/g.eps}
 \caption{Plots of stellar and dark matter misalignments, ellipticities and composed plot.}
 \label{fig:environmentmisalign}
 \end{center}
\end{figure}

\section{Conclusion}
\begin{itemize}
\item Recapitulate what were the main results
\item Answer whether this method is promising
\item What is needed to understand such misalignments better (-> Surely need to understand environment effects better, ...)
\end{itemize}

\section{Acknowledgements}
Acknowledge Dominik Leier...
JIR would like to acknowledge support from SNF grant PP00P2\_128540/1.

\bibliographystyle{mn2e}
\bibliography{paper.bib}

>>>>

% Open Tasks
%
% --- Keywords ---
% - Add keaywords
%
% --- Introduction ---
% - Citiations needed in parametric vs. nonparametric comparison part
% - Citiations needed on effect of environment
% - Section on how this work might be able to rule out certain alternative gravity models
% - Describe theoretical expectations (ellipticities of dm halo and luminous part, misalignments, effect of environment)
% - ...
%
%
% --- Data ---
% - More detailed description how the baryionic mass distribution was measured
% - ...
%
%
% --- Method ---
% - Maybe more details on GLASS and how it works (cite Jonathan's upcoming paper)
% - Plot showing how well the shape informations is recovered when applied to fake data
% - ...
%
%
% --- Results ---
% - Interpretation of 
% - Interpretation of ellipticities plot (why it is relevant that dm halos are significantly rounder)
% - Interpretation of misalignment versus ellipticity plot ('forbidden region', alignment with semi-minor axis, very elliptical galaxies well aligned)

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{myaasmacros}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amsmath}
\usepackage{multirow}

% Some definitions of things I always use here:
\newcommand\bcite[1]{\citeauthor{#1} \citeyear{#1}}

@@

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Somewhere in France, France
}

@@

\begin{abstract}
Understanding the structure and properties of galaxies today is essential to understand how galaxies formed. This poses a larger problem to us than one might imagine. Especially constraining the dark matter distribution in galaxy halos is not trivial. There are only a few methods to directly estimate this distribution. One of them is strong gravitational lensing, an effect caused by all of mass. This phenomenon allows constraining the gravitational potential of lens galaxies and thus the mass distribution. With knowledge of the light distribution it is possible to disentangle the dark matter and baryonic mass components and study their properties.
We use the newly-developped nonparametric framework \textit{GLASS} to model the mass distribution of a total of 15 strong gravitational lensing galaxies, most of which are early-type galaxies, for which we have information on the baryonic mass distribution. We are thus able to separate the mass components. By using a simple shape measure, namely the inverse of the moment of inertia tensor, we can analyze the shapes of the distributions. Analyzing the alignments between the semi-major axes of the distributions, we find that there can be significant misalignments and also cases of alignment between the semi-minor and semi-major axes. Mass thus need not follow light. Furthermore, the ellipticities of the dark matter distributions are always smaller than the baryonic mass distributions. Dark matter halos tend to be more spherical than the light distribution of the galaxy. We also consider the environment of the lenses. We find that lenses in denser environments tend to be more misaligned with either the semi-major or semi-minor axis. Additionally, the semi-major axis of lenses lying in groups or clusters of the dark matter halo might be oriented in direction of the centroid.
\end{abstract}

\begin{keywords}
asdf....
\end{keywords}

\section{Introduction}\label{sec:introduction}

@@

Gravitational lensing is one of the few phenomena that allow probing the matter distribution directly. It makes no distinction between baryonic and dark matter, both components act on light rays in the same way. Gravitational lensing is therefore somewhat special as it allows to \textit{see} dark matter, if we are able to separate the two components. Therefore studying these effects allows us to probe dark matter and its properties. We are typically interested in measuring its distribution, how it clusters, its connection to structure formation, and of course its nature. With strong gravitational lensing galaxies we are effectively probing the matter distribution on relatively small scales. Thus, much insight could be gained in how galaxies form and how they are structured.

The main challenges are however first, estimating the baryonic mass distribution such that the two components can be disentangled (see e.g. \cite{leier11phd}), and second, modelling the lens, i.e. the mass distribution. This work focuses purely on the second aspect. There are in general two classes of methods to reconstruct gravitational lenses, parametric and nonparametric methods. A comparison of these two classes of methods can be found in \cite{2008ApJ...681..814C}. Parametric methods assume a distribution, which depends only on few parameters, such that the problem is well constrained and a best-fitting model can be found. They are computationally cheap, but on the other hand the assumption of a distribution might bias the result (CITATION NEEDED!). Nonparamatric (or free-form) methods assume a set of basis functions and find valid solutions in terms of linear superpositions of these functions. The problem is largely underconstrained. Due to the fact that there is no a priori reason to prefer specific mass models, given that they all satisfy certain prior conditions and reproduce the same gravitational lensing effects, over others, the best-fitting model is of little interest. Much more important is to explore many models and look at the properties of all the models in the ensemble. Typically high-dimensional parameter spaces, which describe all allowed models, need to be sampled in a uniform and uncorrelated way to generate a representative ensemble of solutions. The computational cost of these methods is usually much higher \citep{2008ApJ...679...17C,2012MNRAS.425.3077L}.

With these tools, we can start answering the long-standing question whether mass follows light. It is a well-known fact that for observations to match for example rotation curves of galaxies, the radial density profiles of baryonic mass and dark matter differ. The question is however whether baryonic mass and dark matter need to be aligned and the ellipticities need to agree. If this is not the case, then this needs to be explained by mechanism and effects that produce and sustain these different properties. Especially the environment of lenses might play a significant role (CITATION NEEDED!). It is furthermore unclear whether modified gravity theories could explain possible misalignments between mass and light.

There were already attempts to analyze the alignment of light and mass. \cite{1997ApJ...482..604K} use Singular Isothermal Spheres (SIS) with an external shear term and Singular Isothermal Ellipsoids (SIE) (e.g. \cite{1994A&A...284..285K}) without external shear to model the mass distribution. Applying this to a sample of 17 lenses, they find a strong correlation between light and dark matter distributions, if there is no strong external perturbation, with only a small scatter of less than $10^{\circ}$. They find however no correlation between the axis ratios of the distributions.

Later on, \cite{2006ApJ...649..599K} and \cite{2012ApJ...761..170G} analyzing lens samples of similar sizes and \cite{2009ApJ...690..670T} analyzing a lens sample containing even 70 galaxies also looked for misalignments. They all use SIEs to model the mass distributions. Only \cite{2012ApJ...761..170G} however include an external shear contribution in case more information is available. They neglect the external shear although \cite{1998ApJ...509..561K} find that SIEs with an independent external shear term fit the data better. They find reasonable alignments, the scatter is however larger than the one found in \cite{1997ApJ...482..604K}. \cite{2012ApJ...761..170G} find a rms scatter of $25^{\circ}$ for the whole sample. Even after dropping the almost spherical lenses, the rms scatter is still $18^{\circ}$. They attribute this to a significant external shear contribution. \cite{2009ApJ...690..670T} even find a significant correlation between the misalignment and the local and global environment. In general, lenses in overdense environments seem to be more likely to be misaligned. \cite{2009ApJ...690..670T} and \cite{2012ApJ...761..170G} also find a correlation with a large scatter between the ellipticities of the light and dark matter distributions.

Contrary to earlier approaches, we choose to use a non-parametric modelling technique. We use a pixel-based (see e.g. \cite{1997MNRAS.292..148S}) lens reconstruction framework called \textit{GLASS} \citep{2012MNRAS.425.3077L}. We apply this successfully on a sample of 15 strong gravitational lensing galaxies using stellar mass data of \cite{leier11phd}. Thus, we can disentangle dark and luminous matter and look at the shapes and alignments of these distributions. We can also look at the ellipticities of the reconstructed lenses and compare them to the ones of the stellar distribution. Furthermore, we discuss the environments of these lenses, as they might affect some properties of gravitational lenses in a significant way \citep{2006ApJ...641..169M,2011ApJ...726...84W}. We think that we have enough statistics to make claims about misalignments between the dark matter halo and the stellar mass distribution, and how the halo is oriented compared to the environment.

In Section \ref{sec:data} we present the lens sample and important properties of each lens. In Section \ref{sec:method} we present the reconstruction method we use and our shape measure. Section \ref{sec:results} consists of the results of this analysis. We show the relations we find between misalignment, ellipticities, and the environment.

\section{Data}\label{sec:data}
\begin{itemize}
\item SPS (Ignacio Ferreras, Dominik, etc. \citep{2011ApJ...740...97L})
\item Write why specifically these lenses were chosen, where they're from
\end{itemize}

\begin{table*}
 \begin{center}
  \begin{tabular}{l c c l | l c c l}
   Lens & $z_{L}$ & $z_{S}$ & Type & Lens & $z_{L}$ & $z_{S}$ & Type\\ \hline \hline
   Q0047 & 0.48 & 3.60 & Quad & HE1104 & 0.73 & 2.32 & Double\\
   Q0142 & 0.49 & 2.72 & Double & PG1115 & 0.31 & 1.72 & Quad\\
   MG0414 & 0.96 & 2.64 & Quad & B1152 & 0.44 & 1.02 & Double\\
   B0712 & 0.41 & 1.34 & Quad & B1422 & 0.34 & 3.62 & Quad\\
   RXJ0911 & 0.77 & 2.80 & Quad & MG2016 & 1.01 & 3.27 & Quad$^{1}$\\
   BRI0952 & 0.63 & 4.50 & Double & B2045 & 0.87 & 1.28 & Quad\\
   Q0957 & 0.36 & 1.41 & Double & HE2149 & 0.60 & 2.03 & Double\\
   LBQS1009 & 0.87 & 2.74 & Double & Q2237 & 0.04 & 1.69 & Quad\\
   B1030 & 0.60 & 1.54 & Double & & & & \\
%   SBS1520 & 0.72 & 1.86 & Double
%	B1600 & 0.41 & 1.59 & Double\\
  \end{tabular}
  \caption{List of all the lenses in the sample and their properties. Data is taken from \textit{CASTLES} and \cite{leier11phd}. The angles are measured north-through-east. \newline $^{1}$ Two images merge, thus only three images can be resolved.}
  \label{tab:overview}
 \end{center}
\end{table*}

\begin{table*}
 \begin{center}
  \begin{tabular}{l c c c c c c c}
   \multirow{2}{*}{Lens} & \multirow{2}{*}{Environment$^{a}$} & \multicolumn{2}{c}{Group/Cluster centroid$^{b}$} & & \multicolumn{2}{c}{Other galaxy} & \multirow{2}{*}{References$^{c}$}\\ \cline{3-4} \cline{6-7}
    & & $\Delta\alpha$ [''] & $\Delta\delta$ [''] & & $\Delta\alpha$ [''] & $\Delta\delta$ ['']\\ \hline \hline
   Q0047 & G(9) & $3.6\pm15.6$ & $43.2\pm28.8$ &   & ... & ... & 1\\
   MG0414 & 2 & ... & ... &   & $-0.4$ & $1.5$ & 2\\
   B0712 & 2 & ... & ... &   & $47$ & $-91$ & 3\\
   RXJ0911 & C & $-12.0$ & $-39.8$ &    & ... & ... & 4\\
   Q0957 & C & ... & ... &   & ... & ... & 5\\
   B1030 & 2 & ... & ... &   & $-9.8$ & $8.5$ & 6\\
   PG1115 & G(13) & $-10.8\pm21.0$ & $-3.6\pm15.6$ &   & ... & ... & 1\\
   B1422 & G(17) & $61.2\pm25.2$ & $-10.8\pm23.4$ &   & ... & ... & 1\\
%   SBS1520 & G(5) & ... & ... &   & ... & ... & 7\\
%   B1600 & G(17) & $-8.4^{d}$ & $-1.5^{d}$ &   & ... & ... & 8\\
   B1608 & G(8) & $-0.7$ & $8.0$ &   & ... & ... & 9\\
   MG2016 & C(69) & ... & ... &   & ... & ... & 10\\
   HE2149 & G & $-36.6$ & $14.2$ &   & ... & ... & 11\\
  \end{tabular}
  \caption{List of lenses lying in a group or cluster environment, or that have a companion. \newline $^{a}$ G stands for group, C for cluster, and 2 means that the lens has a companion galaxy. If known, the number of member galaxies including the lens is given. \newline $^{b}$ If not stated differently, the centroid positions are luminosity-weighted estimates. \newline $^{c}$ References for the centroid estimates we used. \newline $^{d}$ Only number-weighted estimates are available. \newline 1: \cite{2011ApJ...726...84W}; 2: \cite{1993AJ....105....1S}; 3: \cite{2002AJ....123..627F}; 4: \cite{2001ApJ...555....1M}; 5: \cite{1998ApJ...504..661C}; 6: \cite{2000ApJ...536..584L}; 7: \cite{2008ApJ...673..778A}; 8: \cite{2007AJ....134..668A}; 9: \cite{2003MNRAS.344..337T}; 10: \cite{2006ApJ...646...85W}; 11: \cite{2006ApJ...646...85W}}
  \label{tab:environment}
 \end{center}
\end{table*}

\begin{table*}
 \begin{center}
  \begin{tabular}{l c c c c}
   Lens & \multicolumn{3}{c}{Time delays [d]} & References\\ \hline \hline
   RXJ0911 & $\Delta t_{BA}=146\pm4$ & ... & ... & 1\\
   Q0957 & $\Delta t_{BA}=416.5\pm1.0^{a}$ & $\Delta t_{BA}=420.6\pm1.9^{b}$ & ... & 2\\
   HE1104 & $\Delta t_{BA}=162.2^{+6.3}_{-5.9}$ & ...  & ...  & 3\\
   PG1115 & $\Delta t_{BA}=12.0^{+2.4}_{-2.0}$ & $\Delta t_{DC}=4.4^{+3.2}_{-2.4}$ & ... & 4\\
%   SBS1520 & $\Delta t_{BA}=125.8\pm2.1$ & ... & ... & 5\\
%   B1600 & $\Delta t_{BA}=51\pm2$ & ... & ... & 6\\
   B1608 & $\Delta t_{BA}=31.5^{+2}_{-1}$ & $\Delta t_{CA}=36.0\pm1.5$ & $\Delta t_{DA}=77.0^{+2}_{-1}$ & 7\\
   HE2149 & $\Delta t_{BA}=103\pm12$ & ... & ... & 8\\
  \end{tabular}
  \caption{Lenses with measured time delays. $\Delta t_{BA}$ corresponds to the time delay between the images A and B with A leading. \newline $^{a}$ measured in the g-band \newline $^{b}$ measured in the r-band \newline 1: \cite{2002ApJ...572L..11H}; 2: \cite{2012A&A...540A.132S}; 3: \cite{2008ApJ...676...80M}; 4: \cite{2010MNRAS.406.2764T}; 5: \cite{2011A&A...536A..44E}; 6: \cite{2000ApJ...544..117B}; 7: \cite{2002ApJ...581..823F}; 8: \cite{2002A&A...383...71B}}
  \label{tab:timedelays}
 \end{center}
\end{table*}

@@

Our sample originally consisted of 21 galaxies. The sample is identical to the one as in \cite{leier11phd} and \cite{2011ApJ...740...97L}. As the lenses \textit{HS0818+1227}, \textit{SBS1520+530}, \textit{B1600+434}, \textit{B1608+656} proved too difficult to reconstruct, we dropped them from this analysis. The final sample consists thus of 17 galaxies. Most of these lenses are early-type galaxies except \textit{B1030+074}, \textit{B1152+200}, and \textit{Q2237+030}. All lenses can also be found in the gravitational lens database \textit{CASTLES}. Characteristics and references of each lens in the sample are listed below. Table \ref{tab:overview} contains all the redshifts and the type of the lenses, Table \ref{tab:environment} lenses with environment information, and Table \ref{tab:timedelays} lenses with measured time delays.

@@

%\textit{B1608+656} (hereafter \textit{B1608}) consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment and the mass distribution along the line of sight have been analyzed by \cite{2006ApJ...642...30F}. They find a group with 8 resp. 9 members if the merging galaxies are counted as 2. No significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. Along the line of sight seem to be four other groups.

@@

\section{Method}\label{sec:method}
\subsection{GLASS}
\begin{itemize}
\item Explain GLASS briefly and link to fake data tests
\end{itemize}

@@

We use the new free-form lens modelling framework \textit{GLASS} \citep{2012MNRAS.425.3077L}. As its predecessor \textit{PixeLens} \citep{2004AJ....127.2604S}, it generates a pixelated mass map of a lens reproducing the image positions perfectly. Using pixels the arrival time surface for a source at position $\boldsymbol\theta_{S}$, taking the geometric and Shapiro contributions into account, can be expressed in terms of the surface mass density $\kappa_{n}$ of the n-th pixel rescaled by the critical density $\Sigma_{crit}$ and the contribution to the gravitational potential by this pixel $\psi_{n}$ \citep{1997MNRAS.292..148S}
\begin{equation}\label{eq:arrival}
\tau(\boldsymbol\theta)=\frac{1}{2}|\boldsymbol\theta |^{2}-\boldsymbol\theta\cdot\boldsymbol\theta_{S}-\displaystyle\sum\limits_{n} \kappa_{n}\psi_{n}(\boldsymbol\theta).
\end{equation}

According to Fermat's principle, lensed images appear where the arrival time surface takes on extremal values $\nabla\tau(\boldsymbol\theta)=0$. Thus, the image positions fulfil
\begin{equation}
\boldsymbol\theta-\boldsymbol\theta_{S}-\displaystyle\sum\limits_{n} \kappa_{n}\alpha_{n}=0,
\end{equation}
where $\alpha_{n}$ is the contribution to the bending angle of the n-th pixel. For pixels, $\alpha_{n}$ and also $\psi_{n}$ can be described by analytic formulae \citep{1997MNRAS.292..148S}. Reconstructing the lens, i.e. describing the surface mass density in every pixel, reduces thus to finding sets of coefficients $\left\{\kappa_{n}\right\}$ of these functions. As the number of pixels sets the resolution, this problem is however highly underconstrained. The number of pixels is typically $\gtrsim100$. The solution to this is noticing that the equation is linear in $\kappa_{n}$. Additional constraints on the lens such as individual time delays between images and their order and all the priors we impose are also linear. The solution space is thus a high-dimensional polytope.

A difficulty in mass modelling is the nonuniqueness of solutions. They are not only not unique, every configuration which fulfils the equation is also equally probable. Therefore, this polytope needs to be sampled in a uniform way to analyze the full range of solutions. has Due to the high dimensionality, sophisticated sampling methods of such volumes are required. Here \textit{GLASS} and \textit{PixeLens} differ in the method they use. \cite{2008ApJ...679...17C} and \cite{2012MNRAS.425.3077L} show that \textit{PixeLens} does not produce uncorrelated random samples. They propose a new sampling algorithm. They prove that generic high-dimensional polytopes indeed are sampled in a much more uniform way. Models in \textit{GLASS} are generated using a MCMC-method with a Metropolis-Hastings algorithm. A new step $x_{i+1}$ of the Markov chain is accepted with the probability
\begin{equation}
 \alpha = \mathrm{min}\left(1,\frac{P(x_{i+1})Q(x_{i},x_{i+1})}{P(x_{i})Q(x_{i+1},x_{i})}\right),
\end{equation}
where $P(x)$ is the probability distribution samples are drawn from, and $Q(x_{i},x_{i+1})$ is the proposal density. It is special in mass modelling that all the valid mass configurations have the same probability, thus $P$ is a constant $\equiv1$. For Markov chains $Q$ has to be symmetric. It is usually taken to be a multivariate Gaussian $\mathcal{N}(x,\hat{\boldsymbol\Sigma})$, where $\hat{\boldsymbol\Sigma}$ is an estimate of the true covariance matrix $\boldsymbol\Sigma$. In principle it could be estimated by an adaptive chain during the random walk, but then the chain is no longer Markovian as it loses the reversibility property. Therefore, first an initial adaptive burn-in phase is run to get an good estimate $\hat{\boldsymbol\Sigma}$, before the real MCMC random walk. Many models can thus be generated and the solution space sampled. The estimation of the covariance matrix is refined in \textit{GLASS} such that models are much less correlated. Further details and tests of the method can be found in \cite{2012MNRAS.425.3077L}.

\subsection{Priors and assumptions}
We use different priors and assumptions to eliminate unphysical mass reconstructions of the lens. First of all, the densities $\kappa_{n}$ need to be always positive. The density must always increase towards the center, why we chose to require the local density gradient in every pixel except the central one to point within an angle of $50^{\circ}$ towards the center. Also, the radial density profile has to decrease monotonically.

To avoid large fluctuations in the density, we choose to add a smoothing prior. The density in any pixel except the central one is at most twice the average density of its neighboring pixels. If there is reason to expect an invariant distribution under a rotation by $180^{\circ}$ (e.g. for doubles), we impose such a symmetry prior.

We also choose to include external shear. Thus, there is an additional term
\begin{equation}
  \gamma_{1}\left(\theta_{x}^{2}-\theta_{y}^{2}\right)+2\gamma_{2}\theta_{x}\theta_{y}
\end{equation}
in (\ref{eq:arrival}) and the subsequent equations. We allow $\gamma_{1}$ and $\gamma_{2}$ to vary, while fixing the maximum absolute value of the shear parameters. For doubles we fix the absolute value to be at most $10^{-2}$ (resp. $10^{-1}$ if no model could be generated for lower values).

We further assume that we know the positions of the images and the lens perfectly. As the range of possible models is much larger than the errors on these quantities \citep{2006ApJ...650L..17S}, this assumption is justified. We also fix a time order of the images. Furthermore, we chose a flat concordance cosmology with $\Omega_{m}=0.28$, $\Omega_{\Lambda}=0.72$, and $H_{0}^{-1}=13.7$.

\subsection{Shape measure}

To get a rough estimate of the shape of the surface mass distribution we use the inverse of the moment of inertia tensor of the mass map. Its eigenvectors and eigenvalues track the shape of the mass distribution. There are thus two quantities which define the shape, the ratio of the eigenvalues $\lambda_{1}/\lambda_{2}$, which describes the ellipticity, and the angle of the semi-major axis $\theta$. This seemingly crude measure was tested on fake data and found to be working well.

Figure \ref{fig:exampleplots} shows example plots of the reconstruction of the lens \textit{PG1115}. They illustrate this shape measure. In this case, there is a significant misalignment between the two mass distributions. Also the uncertainties on the angle estimate are shown. Errors are purely statistical. Figure \ref{fig:exampleplots} also shows the arrival time surface and total, dark, and baryonic mass distributions.

\section{Results}\label{sec:results}
\begin{itemize}
\item Show some examples (stellar and dm contour plots, image of lens, arrival time surface, included mass)
\item Show money plots
\item Comment on a few lenses special lenses
\item Environmental effect?
\end{itemize}
